---
title: "Week 2 - Getting Started"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
library(datasets)

options(digits = 2)
```

# Intro

This week we learn about some basic, but important features in R. This is a R markdown (.Rmd). 
R markdowns are great for reproducibility, as we can explain and document our work using notes, comments, 
code, and outputs. R markdowns can be knitted into html files which are 
easier to follow compared to an R script (e.g., `Week2_GetStarted.R`).

Use the `.Rmd` files during the workshops to follow along and practise coding, and read `.html` files independently in your own time.

# Operators
R's operators are similar to a calculator and other programming languages.

## Arithmetic Operators

```{r}
1+1 # sum
1-1 # subtraction
2*2 # multiplication
2/2 # division
2^2 # power
```

## Logical Operators

```{r operators2}
2*2 == 2*3 # equal
2*2 != 2*3 # not equal
2*2 > 2*4 # larger
2*2 < 2*4 # smaller
2*2 <= 2*4 # smaller or equal to
2*2 >= 2*4 # larger or equal to
``` 

**Do You Know**

You might have noticed that we used `==` to check whether two operands are equal. Why not `=`? 
In R and many programming languages, `=` is an assignment. Specifically in R, 
`=` works similarly to `<-`. They both assign the right operand to the left operand. 

**You Try It**

Check whether the product of 13 and 3 is larger than the product of 29 and 2.

```{r}

```

# Data types
`R` has 5 basic atomic classes:

 * logical, e.g., TRUE, FALSE
 * integer, e.g., 1L, 2L
 * number/real/numeric, e.g., 2 or 2.2
 * character/string/text, e.g.,"a", "b"
 * complex, e.g, 1 + 2i

We can check what data it is, using `typeof()` or `class()` function.

```{r datatype}
typeof(1L)
typeof("PSY3062")
```

# Data structures
## Variables 
Everything in `R` is an object. We create new object by assigning a value using "<-" operator.
Therefore, a `variable` is also an object. When you create an object, the object will appear in 
the environment pane of RStudio (the top right one). We can also perform a task on an object by referring to it by name.

**Tips** 

Every time you create an object, check the environment pane to make sure that it has been successful.

```{r vars}
myclass <- "PSY3062"
```

**You Try It**

Can you see `myclass` in your environment? Check what data type `myclass` is.

```{r }

```

## Vectors
Vector is the most common data structure in `R`.
We create vectors using the `c()` function. You can think of it as "combine".

`x` is a numeric vector

```{r vectors1}
x <- c(13, 22, 3)
```

`y` is a integer vector

```{r integer}
y <- c(1L, 2L, 3L)
```

`z` is logical vector

```{r logical}
z <- c(TRUE, TRUE, FALSE, FALSE)
```

`w` is character vector

```{r character}
w <- c("Monica", "Rachel", "Phoebe", "Chandler", "Joey", "Ross")
```

Examine your vector

```{r vectors2}
typeof(w)
class(w)
```

You can create vectors as sequence of numbers.

```{r seqvector}
series <- c(1:10)
series
```

**You Try It**

Create a new variable call `j` and assign 5 numbers to it.

```{r}

```

Missing values in R are denoted by `NA` and/or `NaN`.

```{r na}
l <- c(1, 2, NA, 4, 5)
is.na(l)
```

## Data frames
A data frame is a very important data structure in R. It is the default data structure in R and is commonly use for statistics.

Let's explore dataset `mtcars`, which is a built in dataset in R. 
Here is the code to load it in our R environment.

```{r dataframe1}
data(mtcars)
```

Check what type `mtcars` is. You can see that it is a data frame.

```{r dataframe2}
class(mtcars)
```

There are several useful functions to explore a data frame.

* `head()` - see first 6 rows
* `tail()` - see last 6 rows
* `nrow()` - number of rows
* `ncol()` - number of columns
* `str()` - display the internal structure of an R object
* `names()` - list the column names

**You Try It**

Explore `mtcars` using the above functions.

 1. How many rows does it have?
 2. How many columns?

```{r}
nrow(mtcars)
```

## Loading data in R
Most of the time we work with a dataset that is our own and not built in R. That means we have to 
load them into R environment.

**Let's load your research report dataset in R.** Recall that everything in R is a object, so is a 
dataset. Therefore, we create a dataset in R using `<-` operator. 

Steps:

 * Call your dataset by a name. Here I call it `d`
 * Call `read.csv()` function to read in a .csv file
 * Within the `read.csv()`, specify the file path to your dataset as a "character string"
 * Run code and check if `d` is in your environment
 
 How to get file path:
 
 * On Macs, right click the file > press and hold the Option key > select "Copy as path"
 * On Windows, right click the file > select "Copy as path" **OR** 
 visit this link https://youtu.be/MVoQhYWJuvw for instructions

Your dataset `d` should appear in the environment panel.

```{r}
# example for Macs
# d <- read.csv("/Users/florale/Library/CloudStorage/OneDrive-MonashUniversity/Work/MonashTA/PSY3062_2023/PSY3062 dataset.csv")

# example for Windows
#d <- read.csv("C:/Users/florale/Downloads/PSY3062 dataset.csv")

## check class
class(d)

# load your own dataset
d <- read.csv("Put your file path here")

```

# Functions
Function is a key feature in `R`. Functions usually take in some sort of data structure 
(value, vector, dataframe etc.), perform a specific task, and return a result. 
We call function in R by their name followed by parentheses `function_name(input)`
For example,

* `class()` checks what type an object is
* `sum()` calculates the sum of vector elements
* `mean()` calculates the mean of vector elements

The inputs, called *arguments* or *parameters*, are information passed to a function.
There can be one or multiple arguments within a function, and they are separated by a comma. 

```{r}
# This function rounds a number
round(1.214294254)

# The same function that rounds a number
# but with an additional argument telling it we want 2 digits
round(1.214294254, digits = 2)
```

**Exercise**

1. Calculate the sum of the `j` variable you created previously.
2. Can you calculate the sum of `l` variable too? What do you get?

```{r}

```

There are also functions that don't require any arguments.

```{r}
getwd()
```

We can also create our own function using the `function()` function. 

The below function, `f1()` takes an argument `name`, add it to
a pre specified character string `is my favourite programming language`, 
and return a concatenated string. 

```{r}
f1 <- function(name) {
  paste(name, "is my favourite programming language")
}
f1("R")
```

**You Try It**

Use `f1()` to print "Python is my favourite programming language".

```{r}


```

Let's create another function that sum two numbers. 
We will call it `f2()`.

```{r}
f2 <- function (x, y) {
  x + y
}

f2(13, 29)
```

**You Try It**

Use `f2()` to sum 89 and 9. What do you get?

```{r}


```

You can see that this function works like the `sum()` function.

```{r}
f2(89, 9)
sum(89, 9)
```

When you are not sure about an function's usage, ask `R` using `??`.

```{r}
??sum
```

I also suggest googling function usage. 
I do this all the time. `R` and most packages in `R` have documentations that explain what a 
function does and how to specify it. 
For example, https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/paste 
explains how to specify the `paste()` function that I just used in `f1()`. 
Google is your best friend when it comes to coding. 
`R` is used by a massive community. There's a good chance that whatever 
issues you run into, someone has been there, done that, and had a solution to it.

Also, most of the time we don't have to write our own functions.
There are amazing developers writing functions and putting them in packages that are freely 
available for us to use in `R`. They make our lives easier. Let's talk packages in the next section.

# Packages
`R` and `RStudio` are like phones. Packages are like apps on our phones. Just like how we download
apps from App Store or Google Play Store, we install packages to use in `R`. Packages contain 
functions that do a lot of hard work for us.

Here are some very helpful packages that we will use throughout this semester. We first need to install them.

```{r, eval=FALSE}
install.packages("data.table")
install.packages("psych")
install.packages("car")
install.packages("ggplot2")
install.packages("JWileymisc")
install.packages("extraoperators")
```

You only need to install packages once, unless you delete and re-install R.
These packages are stored somewhere in your computer. 
You need to load them to R every time you open R. 
We will explore functions from these packages in later weeks.

**Tips** 

Every time you open a R script/ R markdown, load the packages using the `library()` function.

```{r, message=FALSE}
library(data.table)
library(psych)
library(car)
library(ggplot2)
library(JWileymisc)
library(extraoperators)
```

# Save your work in R

This is no brainer, but remember to save your work using CTRL + S or Cmd + S.

# Exercises

Use functions you learnt today to explore the PSY3062 research report dataset.

1. What is the structure of this dataset? How many rows/columns are there? 
2. How many participants were included in this study/dataset?
3. How many variables are in there?

```{r}

```

# Extra materials for R and coding
## DataCamp
DataCamp is an interactive learning platform for many programming languages, including R. 
If you are interested in learning a particular aspect of R or an analysis, 
let me know and
Michelle and I can add new chapters/exercises that you can practise coding in your own time. 

## GitHub
GitHub is a code hosting platform and lets us share our work for transparency.
I store all of our materials on Github, here https://github.com/florale/MonashPSY3062R. 
On GitHub, you will be able to see all changes/updates I make to our workshop materials. 
You can download the materials from Moodle or Github. 
We won't cover Github in PSY3062, and you will work with Github more in Honours.
If you are keen to learn, feel free to chat with me about it.