---
title: "Week 4 - Scoring"
author: Your name
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: sandstone
    highlight: zenburn
    css: "florastheme.css"
    toc: yes
    toc_float: yes
    collapsed: no
    smooth_scroll: no
    toc_depth: 4
    fig_width: 6
    fig_height: 4.5
    fig_caption: yes
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
library(datasets)

options(digits = 2)

```

# Intro

This week we talk about how to score data in `R`, using the `data.table` package.


# Packages

We are going to start with loading a few of the packages that you
should have already installed from Week 2.

```{r pkg, message=FALSE}
library(data.table)
library(psych)
library(car)
library(ggplot2)
library(JWileymisc)
library(extraoperators)
```

# Import Data

```{r dat}
d <- fread("/Users/florale/Library/CloudStorage/OneDrive-MonashUniversity/Work/MonashTA/PSY3062_2022/Tutorial 2022/Tutorial 3/Tutorial 3 - Perceived Control of Internal States.csv")
```

# Scoring

When we work with our own data, often we have to score our variables from the individual item scores. 
Scales/subscales in psychology are usually scored by summing or taking the average.

**Tips**

Always refer to the codebook and instructions for scoring. They have information about the scale, 
reverse scoring that are crucial for computing the correct sum or average score.

## Reverse scoring
We often have to reverse some items before computing the sum or average. You learnt about 
reverse scoring in last week's tutorial. We are going to subtract the items that require reverse 
scoring from `6`.

**Exercise**

1. Why `6`?
2. What items need to be reversed?

```{r revscore}
d[ , pc3 := 6 - pc3]
d[ , pc3r := 6 - pc3]

```

## Average score

We calculate average scores using the `rowMeans()` function (take the mean for a row of data).

* `.SD` is a special symbol of `data.table`. It refers to a **S**ubset of **d**ata.table. Its job is to 
  hold the data for the a particular group defined by us.
* `na.rm` stands for **R**e**m**ove missing data (**NA**). In this case, it asks `R` to remove missing 
  data when calculating the mean.
* `.SDcols` specifies the columns of the `data.table` that are included in `.SD`.

```{r mean}
d[, pc_mean := rowMeans(.SD, na.rm = TRUE),
                         .SDcols = c("pc1", "pc2", "pc4")]
```

**You Try It**

Calculate the average of all variables.

```{r}

```

## Total score

There are two ways to calculate the total score:

* The simplest approach is literally adding items using `+`. This works fine if there is no 
  missing data. However, if a participant has missing data on *any* of the items, R will return 
  missing value `NA` for the total score for this participant.
  Recall when this happened to us trying to calculating the sum of 
  the `x` vector that contains a missing value last week.
* Another approach is using `rowMeans()` function. We use `rowMeans()` to get an average score. 
  We then get back to the "sum" scale by multiply the result to the number of items 
  (e.g., 10 if a scale has 10 items). This works well if we have missing value on a small number of 
  items, as we can keep participants with small amount of missing data. This doesn't work if the 
  participants have missing data on all items (obviously).

```{r total}
# add items
d[, pc_total := pc1 + pc2 + pc3]

# create a total score using rowMeans
d[, pc_total := rowMeans(.SD, na.rm = TRUE) * 3,
                         .SDcols = c("pc1", "pc2", "pc3")]
```

**You Try It**

Calculate the sum of all `pc` variables.

```{r}

```

## T-Score

T Scores are calculated via the scoring conversion tables. 

Let's first load the dataset for week 3 tutorial - *Tutorial 3 - T scores.csv*, 
which contains variables that require converting to T-scores.

```{r tscore}
d_tscore <- fread("~/Library/CloudStorage/OneDrive-MonashUniversity/Work/MonashTA/PSY3062_2022/Tutorial 2022/Tutorial 3/Tutorial 3 - T scores.csv")
```

### Step 1 - Calculate raw/total score

We first need the total score. Here is the code to calculate the total scores for only 3 items.
Your code should include all items of the scale.

```{r}
d_tscore[, SS_COMP_Total := rowMeans(.SD, na.rm = TRUE) * 3,
                                     .SDcols = c("SS_COMP_1", "SS_COMP_2", "SS_COMP_3")]
```

### Step 2 - Convert to T-score

**Conversion chart** : https://www.healthmeasures.net/images/promis/manuals/PROMIS_Companionship_Scoring_Manual.pdf

This task is similar to assigning values to a new variable. Here we match values of our raw scores with 
the corresponding T-scores to create a new T-score variable. The code formula is below:

```{r}
# Create the a vector of scale (T-)scores using the conversion chart
T_score <- c(25.2, 29.5, 31.8, 33.8, 35.5, 37.3, 
             39.1, 40.8, 42.5, 44.3, 46.2, 48.1, 
             50.0, 52.1, 54.3, 56.9, 63.1)

# Create a vector of raw scores
# If applicable, use a sequence of numbers
Raw_score <- 4:20

# Convert to t-score
d_tscore[, SS_COMP_TScore := T_score[match(SS_COMP_Total, Raw_score)]]
```

Check your data after recoding using the `View(d_tscore)` function.

**You Try It**
In this week tutorial, calculate the T-score of the entire scale.

# Reliability
Reliability is assessed using Cronbach's alpha. Cronbach's alpha is calculated from the individual 
items, not the total nor mean.

Here is the code to calculate Cronbach's alpha of three first `pc` items. 
Your code should include all items of the scale.

```{r alpha}
## calculate Cronbach's alpha
psych::alpha(as.data.frame(d[, .(pc1, pc2, pc3)]))
```

**You Try It**

Calculate Cronbach's alpha for the entire scale. Pay attention to the warning. 

```{r}
psych::alpha(as.data.frame(d[, .(pc1, pc2, pc3, pc4, pc5, pc6, pc7, pc8, pc9,
                                 pc10, pc11, pc12, pc13, pc14, pc15, pc16, pc17,
                                 pc18)]))
```

Let's run this again using the `check.keys = TRUE` option. This option tells `alpha()` function to 
automatically reverse items when calculating Cronbach's alpha.

```{r}
psych::alpha(as.data.frame(d[, .(pc1, pc2, pc3, pc4, pc5, pc6, pc7, pc8, pc9, pc10,
                                 pc11, pc12, pc13, pc14, pc15, pc16, pc17, pc18)]), check.keys = TRUE)
```

**Interpreting R Output**

You can see that R gives us several outputs. You can learn about interpreting these results by running 
this code `?alpha`. The most helpful output here is the `raw_alpha`. It tells you about the Cronbach's alpha, 
which is `r psych::alpha(as.data.frame(d[, .(pc1, pc2, pc3, pc4, pc5, pc6, pc7, pc8, pc9, pc10, pc11, pc12, pc13, pc14, pc15, pc16, pc17, pc18)]), check.keys = TRUE)$total$raw_alpha`. 

*What does this value indicate? Does this scale have good reliability?*

```{r}
# run ?alpha below if interested

```

# Exercises - Data management for your research project

## Scoring
Score the variables you have chosen for your research report. This should include reverse scoring and 
computing the final score by either summing or averaging individual items.

```{r}

```

## Reliability
Calculate the Cronbach's alpha for the scales you have chosen. Do your scales have good reliability?

```{r}

```