---
title: "Week 12 - Control Flow in R"
date: "`r Sys.Date()`"
output: html_document
---

# Intro

Welcome to the final workshop! As requested, this week is a brief introduction to control flow in `R`. 
We discuss if/else statement, `ifelse()`, `for` loops, and `apply()` family functions

Flow control or conditional execution is critical in a lot of code.
Often some code should only be run under certain conditions and other 
code run under other conditions. There are two primary tools of control flow: choices and loops. 

**Choices**

1. `if...else...`
2. `ifelse()`

**Loops**

3. `for` loops
4. `apply()` family functions

These are a few key topics you should understand and be able to code in `R`,
at least at a basic level.

## Packages and Data

The `flights` dataset contains information about all flights that departed from NYC 
(e.g. EWR, JFK and LGA) to destinations in the United States, Puerto Rico, and the American Virgin Islands 
in 2013: 336,776 flights in total. 

```{r pkg, message=FALSE}
library(data.table)

# data
install.packages("nycflights13")
library(nycflights13)
data("flights")

# make datatable
flights <- as.data.table(flights)
```

## `flights` variables

```{r}
??flights
str(flights)
```

* `year`, `month`, and `day`: Date of departure.
* `dep_time` and `arr_time`: Departure and Arrival time, in Eastern time zone.
* `dep_delay` and `arr_delay`: Departure and Arrival delay, in minutes.
* `sched_dep_time` and `sched_arr_time`: Scheduled Departure and Arrival times.
* `carrier`: Carrier, abbreviated.
* `flight`: Flight number.
* `tailnum`: Tail number of the airplane.
* `origin`: Flight origin, airport code.
* `dest`: Flight destination, airport code.
* `air_time`: Time in the air, in minutes.
* `distance`: Distance between the departure and arrival airports, in miles.
* `hour` and `minute`: Scheduled departure broken into hour and minute.

# `if/else` statements
The basic form of an `if` statement in `R` is as follows:

```{r}
if (condition) true_action
if (condition) true_action else false_action
```

If condition is `TRUE`, `true_action` is evaluated; 
if condition is `FALSE`, the optional `false_action` is evaluated.

**When to use** `if/else` statement

```{r}
a <- 13
b <- 29

if (a > b) {
  print("a is greater than b")
  } else {
    print("b is greater than a")
  }
```

**Example** Evaluating a condition before an assignment.

Let's look at an example of fitting a `lm()` model predicting departure delay `dep_delay` 
by flight origin `origin`. 
In addition, we want to run this model only when our predictor `origin` is continuous.
Note that this is not a very good coding practice, 
as we can simply check the variable type. 
This type of evaluation comes handy in scenarios such as writing functions.


```{r}
if (identical(class(flights$origin), "numeric")) {
  m <- lm(dep_delay ~ origin, data = flights)
  } else {
    stop("predictor should be a numeric variable")
  }

# check variable type
class(flights$origin)
```

# `ifelse()`
`ifelse()` is a vectorised function with test, yes, and no vectors.

```{r, eval = FALSE}
ifelse(test, yes, no)
```

**When to use** `ifelse()`.

```{r}
c <- c(1, 2, 3, 8, 9)
ifelse(c %% 2 == 0, "even", "odd")
```

**Example** Basic data management

Create a new variable `dep_status` to indicate whether the departure time was on-time, early, or delayed 
based on the departure delay variable (`dep_delay`).

```{r}
flights[, dep_status := NA]
flights[, dep_status := ifelse(dep_delay == 0, "On time", dep_status)]
flights[, dep_status := ifelse(dep_delay > 0, "Delayed", dep_status)]
flights[, dep_status := ifelse(dep_delay < 0, "Early", dep_status)]
```


```{r}
table(flights$dep_status)
```

**Example** Advanced data management
 
Create a new variable `status` to indicate whether a flight was on-time, early, or delayed 
based on both departure (`dep_delay`) and arrival time (`arr_delay`).

```{r}
flights[, status := NA]
flights[, status := ifelse(dep_delay == 0 & arr_delay == 0, "On time", status)]

flights[, status := ifelse(dep_delay == 0 & arr_delay > 0, "Delayed Arrival", status)]
flights[, status := ifelse(dep_delay > 0 & arr_delay == 0, "Delayed Departure", status)]

flights[, status := ifelse(dep_delay == 0 & arr_delay < 0, "Early Arrival", status)]
flights[, status := ifelse(dep_delay < 0 & arr_delay == 0, "Early Departure", status)]

flights[, status := ifelse(dep_delay < 0 & arr_delay > 0, "Early Departure and Delayed Arrival", status)]
flights[, status := ifelse(dep_delay > 0 & arr_delay < 0, "Delayed Departure and Early Arrival", status)]

flights[, status := ifelse(dep_delay > 0 & arr_delay > 0, "Delayed Departure and Arrival ", status)]
flights[, status := ifelse(dep_delay < 0 & arr_delay < 0, "Early Departure and Arrival ", status)]
```


```{r}
table(flights$status)
```

# `for` loop
`for` loops are used to iterate over items in a vector. They have the following basic form:

```{r, eval = FALSE}
for (item in vector) perform_action
```

**When to use** `for` loop.

```{r}
for (d in 1:13) {
  e <- d + 29
  
  print(e)
}
```

**Example** Running models with for loop

Let's run a series linear models predicting Distance flown between airports (`distance`) from 
amount of time spent in the air (`air_time`) and `carrier`,
stratified by month of departure. 

These models have the same set of predictors and outcomes, but different datasets. 
Therefore, we loop through 12 datasets (for 12 months) to run our models. 
We can use for loop to run our models on each dataset.

```{r}
results <- list()
for (m in 1:12) {
  model <- lm(distance ~ air_time + factor(carrier), data = flights[month == m])
  
  results[[m]] <- model
}
```

Inspect model results.

```{r}
summary(results[[1]])
```

**Example** Combining for loop and if/else to convert variables into factors.

In this example, we loop through each variable in the `flights` dataset, 
check whether it is a character variable and if so, convert to a factor.

```{r}
# create a copy of the dataset
tmp <- copy(flights)

# convert variables using both for loop and if statement
for (j in colnames(tmp)) {
  if(identical(class(tmp[[j]]), "character")) {
    tmp[, (j) := as.factor(get(j))]
  }
}
```

Let's compare the two datasets.

```{r}
str(flights)
str(tmp)
```

# `apply` Functions
`apply` functions are a family of functions in base `R` which allow you to repetitively perform an 
action on multiple chunks of data. 
`apply` family in r contains `apply()`, `lapply()`, `sapply()`, `mapply()`, and `tapply()`. 
`apply` functions essentially a for loop, but run faster than loops and often require less code, 
because they are vectorised. 
Let's look at the three most common, `apply()`, `lapply()`, and `sapply()`.

| Functions        | Syntax                   |
|:----------------:|:------------------------:|
| `apply`          | apply(X, MARGIN, FUN, …) |
| `lapply`         | lapply(X, FUN, …)        |
| `sapply`         | sapply(mylist, FUN)      |

**When to use** `apply()`

`apply()` function takes a *data frame* or *matrix* as input and 
can be applied by the rows (margin = 1) or by the columns (margin = 2). 
It returns output in *vector*, *list*, or *array.*

```{r}
# first remove missing data
flights <- flights[complete.cases(flights)]

# calculate mean of each numeric variable
get_mean <- apply(flights[, c(4:9)], 2, mean)
get_mean

# calculate SD of each numeric variable
get_sd <- apply(flights[, c(4:9)], 2, sd)
get_sd
```

**When to use** `lapply()` and `sapply()`

The `lapply` function applies a function to a *data frame*, *list* or *vector*, 
and return a *list* of the same length. 
`sapply()` does the same job as `lapply()` function but returns a *vector*, 
instead of a list. **They both don't need MARGIN.**


```{r}
get_lmean <- lapply(flights[, c(4:9)], mean)
get_smean <- sapply(flights[, c(4:9)], mean)
```

**Example** Running multiple models with `apply()`.

Here is an example where we use Flight origin (`origin`) to predict 
Departure and Arrival time (`dep_time` and `arr_time`),
Departure and Arrival delay (`dep_delay` and `arr_delay`), and 
Scheduled departure and Arrival times (`sched_dep_time` and `sched_arr_time`), respectively.

These models have the *same predictor(s)* but *different outcomes*.
Instead of running 6 individual models (one for each outcome), 
we can use either `lapply()` or `sapply()` to essentially loop through our outcome variables.

```{r}
# notes. tmp[, c(4:9)] is a subset of tmp that includes only the 6 outcomes variable

list_of_model <- lapply(tmp[, c(4:9)], function(dv){
  m <- aov(dv ~ origin, data = tmp)
  summary(m)
})

sapply(tmp[, c(4:9)], function(dv){
  m <- aov(dv ~ origin, data = tmp)
  summary(m)
})

```
