---
title: "Week 3 - Working with Data 1 and Scoring"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(datasets)

options(digits = 2)

```

# Intro

This week we talk about how to work with data in data table and score data in `R`. 
As a brief refresher, data frame is the default structure in 
R. `data.table `is an enhanced version of data frame. 
A big advantage of `data.table` is how fast it is, especially with large datasets. 
In this workshop, we will use `data.table` to manage our data using the `data.table` package.


# Starting your R project
When working on a project, we likely create many files for different purposes. For example, 
throughout this series of workshops, we have multiple HTML files, R Markdowns and R Scripts. 
We can manage our files conveniently through `R` Project in RStudio. 

Let's create a R project for the PSY3062 unit. You can save files from the workshops and 
your analysis for the research report in this folder.

Steps to create a `R` project:

* From the menu bar, select `File > New Project...`.
* Select `New Directory` from the popup window.
* Select `New Project`.
* Pick a name for your project.
* Click `Create Project`.
* Move all R files related to PSY3062 to this folder.

For more detailed instructions, visit https://bookdown.org/daniel_dauber_io/r4np_book/starting-your-r-projects.html.

**Tips** 

Always open `R` project everytime you open `R` Studio. You can do this clicking the .Rproj file in 
`R` Studio. This will let RStudio set the working directory, where all of the files you work on are saved to.

# Packages

We are going to start with loading a few of the packages that you
should have already installed from Week 2.

```{r pkg, message=FALSE}
library(data.table)
library(psych)
library(car)
library(ggplot2)
library(JWileymisc)
library(extraoperators)
```

**Tips** 

* If you encounter trouble installing packages, try the following code: 
`install.packages("XXX", dependencies= TRUE)`.
* If you receive an error saying `Error in library(XXX) : there is no package called ‘XXX’`, it means 
you have not installed the package. Running `install.packages("XXX")` then `library(XXX)` should 
fix the problem.

# Loading dataset as data table

We can create a data table when loading a dataset in R in 2 ways:

1.

* read dataset as usual using `read.csv()` function.
* use `as.data.table()` function to make it a data table.

2. use `fread()` function (fast read)

Let's load the dataset we use for *week 3 tutorial*.

```{r data}
d0 <- as.data.table(read.csv("Put your file path to the perceived control csv file here"))
d <- fread("Put your file path to the perceived control csv file here")
```

These methods both work well. However, `fread()` is more speedy, which is helpful when you have a 
large dataset. We can check whether `d` and `d0` are the same dataset using the `identical()` function.

```{r checkd}
identical(d, d0)
```

Let's quickly inspect `d`. You can also view by clicking on `d` in your environment pane

```{r checkd1}
head(d)
```

# Working with Data using data.table
## Data table syntax

In data table, we write our code inside `[...]`. The syntax of data table is `DT[i, j, by]`,
which means subset/reorder rows using `i`, then calculate `j`, grouped by `by`.

* `DT` : name of your data object.
* `i` : where you specify the row (i.e., observations, participants) that you want to consider.
* `j` : what to do with the column (i.e., variables), such as creating or modifying.
* `by` : would you like the task to happen within a group. For example, you might like to perform 
   an function only with females/males.

## Creating a new variable
Columns can be referred to as if they are variables. This means creating a variable is done in `j`. 
In `data.table`, we use a special symbol `:=` for assignment within the `[...]`.

For example, we want to make a new variable `NewVar` that assign the same value of 1 to all participants. 
As this variable applies to all participant, `i` is left blank.
 
```{r create}
d[ , NewVar := 1]
```

**Tips** 

If you receive an error saying `Error in `:=`(NewVar, 1) : could not find function ":="`, that means
you don't have the package that has the function ":=". This function is from the `data.table` package,
so you need to load it using the `library()` function.

**You Try It**

Create a new variable in `d`, call it `Class`, and assign `PSY3062` to it.

```{r}


```

## Dropping/Deleting variable

When we want to delete variables, we assign `NULL` to it. Here is the code to drop variable `NewVar` 
from `d`:

```{r drop}
d[, NewVar := NULL]
```

**You Try It**

Drop the `Class` variable.

```{r}


```

# Subsetting dataset
Subsetting is a useful feature in `R` that can be used to select or filter observations/participants 
and variables. Subsetting can be performed using both rows (represented by `i`) and columns (represented by `j`). 
For example, you can subset the PSY3062 dataset to keep only female participants (subsetting by rows) 
or the three variables that you're using for your research project (subsetting by columns).

## Subsetting rows in `i`
Get the first row from `d`. Key word here is row. 
We need to provide `i` with the row index that we are after. 
This returns a data table d with all columns at the first row.

```{r subset}
d[1, ]
```

**You Try It**

Subset the first 10 rows from `d`. Remember that we are asking for multiple rows. When there are 
multiple values, we use the `c()` function to make it a vector, e.g., `c(1, 2)` not `1, 2`. 
We can also create sequence of numbers using `:`.

```{r}

```

## Subsetting columns in `j`
Get the first column from `d`. Because we are selecting columns, we provide `j` with the column 
indices that we are after.

```{r}
d[ , 1]
```

**You Try It**

Subset columns 1, 2, 3, 8, 9 from `d`.

```{r}

```

## Subsetting data by variable names
We can also subset using the names of the variables. For example, we only want variable `pc1` and `pc2`.
This is helpful when we know the variable names, but we don't know which column indices they are.

```{r}
d[, .(pc1, pc2)]
```

## Create a sub dataset

We can create a new dataset with our chosen variables just like how we create any objects in `R`.

```{r}
d1 <- d[, .(pc1, pc2)]
```

# Missing data
## Checking for missing data
We can check for missing data using the `is.na()` function.

This shows us where pc1 has missing data. We are performing this task in the first section of the
squared brackets `i`, which means that we are subsetting rows .

```{r missing}
d[is.na(pc1)]
```

## Removing missing data
We can subset `d` to include only rows that don't have missing data in `pc1`. 
To do this, we add `!` before `is.na()`. `!` means NOT. Here, we make a new dataset `d2` that only 
includes rows that don't have missing data on `pc1`.

```{r}
d2 <- d[!is.na(pc1)]
```

**Exercise**

How many rows do `d`, `d1`, and `d2` have? How many rows have missing data on `pc1`?

Alternatively, you can use the `complete.cases()` function to find rows that don't have missing data.

```{r}
d3 <- d[complete.cases(pc1)]
```

`complete.cases()` and `!is.na()` work similarly in this example. We can confirm this by checking if 
`d2` and `d3` are identical using `identical()` function.

```{r}
identical(d2, d3)
```

We can also perform this on the entire dataset, rather than just only `pc1`.

```{r}
d4 <- d[complete.cases(d)]
```

# Scoring and Reliability

When we work with our own data, often we have to score our variables from the individual item scores. 
Scales/subscales in psychology are usually scored by summing or taking the average.

**Tips**

Always refer to the codebook and instructions for scoring. They have information about the scale, 
reverse scoring that are crucial for computing the correct sum or average score.

## Reverse scoring
We often have to reverse some items before computing the sum or average. You learnt about 
reverse scoring in last week's tutorial. We are going to subtract the items that require reverse 
scoring from `6`.

**Exercise**

1. Why `6`?
2. What items need to be reversed?

```{r revscore}
d[ , pc3 := 6 - pc3]
```

## Average score

We calculate average scores using the `rowMeans()` function (take the mean for a row of data).

* `.SD` is a special symbol of `data.table`. It refers to a **S**ubset of **d**ata.table. Its job is to 
  hold the data for the a particular group defined by us.
* `na.rm` stands for **R**e**m**ove missing data (**NA**). In this case, it asks `R` to remove missing 
  data when calculating the mean.
* `.SDcols` specifies the columns of the `data.table` that are included in `.SD`.

```{r mean}
d[ , pc_mean := rowMeans(.SD, na.rm = TRUE),
                         .SDcols = c("pc1", "pc2", "pc3")]
```

**You Try It**

Calculate the average of all variables.

```{r}


```

## Total score

There are two ways to calculate the total score:

* The simplest approach is literally adding items using `+`. This works fine if there is no 
  missing data. However, if a participant has missing data on *any* of the items, R will return 
  missing value `NA` for the total score for this participant.
  Recall when this happened to us trying to calculating the sum of 
  the `x` vector that contains a missing value last week.
* Another approach is using `rowMeans()` function. We use `rowMeans()` to get an average score. 
  We then get back to the "sum" scale by multiply the result to the number of items 
  (e.g., 10 if a scale has 10 items). This works well if we have missing value on a small number of 
  items, as we can keep participants with small amount of missing data. This doesn't work if the 
  participants have missing data on all items (obviously).

```{r total}
# add items
d[, pc_total := pc1 + pc2 + pc3]

# create a total score using rowMeans
d[, pc_total := rowMeans(.SD, na.rm = TRUE) * 3,
                         .SDcols = c("pc1", "pc2", "pc3")]
```

**You Try It**

Calculate the sum of all `pc` variables.

```{r}


```

## Reliability
Reliability is assessed using Cronbach's alpha. Cronbach's alpha is calculated from the individual 
items, not the total nor mean.

```{r alpha}
## calculate Cronbach's alpha
psych::alpha(as.data.frame(d[, .(pc1, pc2, pc3)]))
```

**You Try It**

Calculate Cronbach's alpha for the this scale. Pay attention to the warning. 

Let's run this again using the `check.keys = TRUE` option. This option tells `alpha()` function to 
automatically reverse items when calculating Cronbach's alpha.

```{r}
psych::alpha(as.data.frame(d[, .(pc1, pc2, pc3, pc4, pc5, pc6, pc7, pc8, pc9, pc10,
                                 pc11, pc12, pc13, pc14, pc15, pc16, pc17, pc18)]), check.keys = TRUE)
```

**Interpreting R Output**

You can see that R gives us several outputs. You can learn about interpreting these results by running 
this code `?alpha`. The most helpful output here is the `raw_alpha`. It tells you about the Cronbach's alpha, 
which is `r psych::alpha(as.data.frame(d[, .(pc1, pc2, pc3, pc4, pc5, pc6, pc7, pc8, pc9, pc10, pc11, pc12, pc13, pc14, pc15, pc16, pc17, pc18)]), check.keys = TRUE)$total$raw_alpha`. 

*What does this value indicate? Does this scale has good reliability?*

```{r}
# run ?alpha below if interested

```

## T-Score

T Scores are calculated via the scoring conversion tables. 

Let's first load the dataset for  *week 3 tutorial*, which contains variables that require 
converting to T-scores.

```{r tscore}
d_tscore <- fread("put your file patht to the t-score csv file here")
```

### Step 1 - Calculate raw/total score

We first need the total score. Here is the code to calculate the total scores for only 3 items.
Your code should include all items of the scale.

```{r}
d_tscore[, SS_COMP_Total := rowMeans(.SD, na.rm = TRUE) * 3,
                                     .SDcols = c("SS_COMP_1", "SS_COMP_2", "SS_COMP_3")]
```

### Step 2 - Convert to T-score

**Conversion chart** : https://www.healthmeasures.net/images/promis/manuals/PROMIS_Companionship_Scoring_Manual.pdf

This task is similar to assigning values to a new variable. Here we match values of our raw scores to 
the corresponding T-scores to create a new T-score variable. The code formula is below:

```{r}
# Create the a vector of scale (T-)scores using the conversion chart
T_score <- c(25.2, 29.5, 31.8, 33.8, 35.5, 37.3, 
             39.1, 40.8, 42.5, 44.3, 46.2, 48.1, 
             50.0, 52.1, 54.3, 56.9, 63.1)

# Create a vector of raw scores
# If applicable, use a sequence of numbers
Raw_score <- 4:20

# Convert to t-score
d_tscore[, SS_COMP_TScore := T_score[match(SS_COMP_Total, Raw_score)]]
```

Check your data after recoding using the `View(d_tscore)` function.

**You Try It**

In this week tutorial, calculate the T-score of the entire scale.

```{r}


```

# Exercises - Data management for your research project
## Task 1 - Loading data

Write the code that loads your research report dataset in the code chunk below.

**Hints**

Revisit `Loading data in R` section of our Week 2 workshop for instructions.

```{r}

```

## Task 2 - Subsetting your variables
Create a new dataset by subsetting only the variables you plan to use for your research report.

**Hints**

Revisit `Subsetting columns` section in this week workshop for instructions. You might like to specify 
the columns/variables using their names rather than the column indices.

```{r}

```

## Task 3 - Removing missing data

Create a new dataset that removes missing data from your research report dataset in the code chunk below.

**Hints**

Use either `!is.na()` or `complete.cases()` function. Always inspect your dataset after running the 
code to make sure it works as intended.

```{r}

```

## Task 4 - Scoring
Score the variables you have chosen for your research report. This should include reverse scoring and 
computing the final score by either summing or averaging individual items.

```{r}

```

## Task 5 - Reliability
Calculate the Cronbach's alpha for the scales you have chosen. Do your scales have good reliability?

```{r}

```


