---
title: "Week 9 - Factorial ANOVA"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: sandstone
    highlight: zenburn
    toc: yes
    toc_float: yes
    collapsed: no
    smooth_scroll: no
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    fig_caption: yes
    number_sections: true
---
# Intro

This week we learn how to conduct factorial analysis of variance (ANOVA), 
specifically two-way ANOVA (ANOVA with two factors).

# Packages

```{r pkg}
library(data.table)
library(psych)
library(car)
library(ggplot2)
library(JWileymisc)
library(extraoperators)
library(lm.beta)
library(visreg)
library(olsrr)
library(corrplot)
library(ggsci)
library(emmeans)
library(effectsize)
library(onewaytests)
```

A new package (again) that helps with getting descriptives by groups.

```{r}
install.packages("dplyr")
library(dplyr)
```

# Data

```{r}
d <- fread("week6_dataset.csv")
```

**Research Questions**
We are going to use ANOVA to address the following question: 
* Does Variable1 vary according to Alcohol Risk and Sex?

## Data management
Check data types of the main variables. 

```{r}
is.numeric(d$Variable1)
is.factor(d$AUDITCat)
is.factor(d$Sex)
```

Again, as we just reloaded our dataset, `AUDITCat` and `Sex` are not factors yet. 
Now quickly convert them.

```{r}
d[, AUDITCat := as.factor(AUDITCat)]

d[, Sex := factor(
  Sex,
  levels = c(0,1),
  labels = c("Male", "Female"))]
```

Check data type again to make sure it works.

```{r}
class(d$AUDITCat)
class(d$Sex)
```

# Descriptive Statistics
We are going to use package `dplyr` to get our descriptive statistics.

```{r}
d %>%
  group_by(AUDITCat, Sex) %>%
  summarise(
    count = n(),
    mean = mean(Variable1, na.rm = TRUE),
    sd = sd(Variable1, na.rm = TRUE)
  )
```

# ANOVA Model

As usual, `aov()` function can be used to fit an ANOVA model to answer our research question: 
*Are there sex differences in Variable1 across Alcohol Risk Categories?*

A two-way ANOVA requires us to include an interaction term between the two categorical variables.

**Exercise**
* Run a ANOVA model called `m0` using `aov()` function that include an interaction term 
  between `AUDITCat` and `Sex`. 
* Run a LM model called `m` using `lm()` function that include an interaction term 
  between `AUDITCat` and `Sex`. 

```{r}
m0 <- 
m <- 
```

Summarise our model results. 

```{r}
summary(m0)
```


```{r}
Anova(m)
```

## Understanding and Interpreting Two-way ANOVA
We can treat an interaction ANOVA model just like how we do with moderated regression.

* For non-sig model, it's a common practice to drop the interaction and run a normal model 
  to test the main effects of our predictors.
* For sig model, we interpret the interaction effect, conduct pairwise comparison, and graph the results.

**EFFECTS OF THE DAY**

* **Main Effects** The effect of one predictor, ignoring (controlling for) the other predictor.
* **Simple Effects** The effect of one predictor, conditioned on the level of another predictor (usually 0). 
* **Interaction Effects**  The effect of one predictor, dependent on the level of another predictor. 

### Non-significant two-way ANOVA
As usual, we drop the interaction and run multiple regression/ANOVA model.

```{r}
mns <- lm(Variable1 ~ AUDITCat + Sex, data = d)
summary(mns)
```

We can write up this model like a non-sig moderated regression/ANOVA. 
**See week 7 and 8 for sample write ups.**

### Significant two-way ANOVA
When our model is significant, we interpret interaction effect. 
A significant p-value in an ANOVA test shows that some of the group means differ, 
but we don’t know which pairings of groups differ. 
*Multiple* pairwise comparisons can be performed to see if the mean differences 
between certain pairs of groups are statistically significant.

#### Pairwise comparison
##### Pairwise comparison for the interaction effect.
First, we obtain the estimated marginal means using `emmeans()`.

```{r}
em <- emmeans(m, c("AUDITCat", "Sex"), adjust = "bonferroni")
```

We can then obtain our multiple pairwise comparisons by summarising the results. 

```{r}
summary(em , infer = TRUE)
```

We can get all simple main-effect comparisons using the code below. 
We get two contrasts, one for Alcohol Consumption and one for Sex.

```{r}
contrast(em, "consec", simple = "each")
```

##### Pairwise comparison for the main effects.
Works the same as post-hoc analyses for one-way ANOVA.

* Main effect of Alcohol Consumption

```{r}
emmeans(m, pairwise ~ AUDITCat)
```

* Main effect of Sex

```{r}
emmeans(m, pairwise ~ Sex)
```

#### Visualising resutls
As usual, `visreg` can be used to graph the results from the `lm` model. 
Here, we separate our results by males and females.

```{r}
visreg(m,
       xvar = "AUDITCat", by = "Sex",
       overlay = FALSE, partial = FALSE, band = TRUE, rug = FALSE, gg = TRUE) +
  ylab("Variable1") + xlab("Alcohol Consumption")
```

Alternatively, we can use the `emmip()` function to graph our result. 
We now separate our results by levels of alcohol consumption

```{r}
emmip(em, ~ Sex | AUDITCat, CIs = TRUE) +
  ylab("Variable1") + xlab("Alcohol Consumption")
```
